\documentclass{article}

\input{D:/Notes/others/myHeadings.tex}

\title{HW2}
\author{Yuanxing Cheng, A20453410, MMAE-501-f22}

\begin{document}

\maketitle

\section*{1}

\begin{myleftlinebox}
    Determine those values of \(\lambda\) for which the following set of equations may possess a nontrivial solution
    \[
        \begin{cases}
            3 x_1+x_2-\lambda x_3 = 0\\
            4 x_1 -2x_2- 3 x_3 = 0\\
            2\lambda x_1+4x_2+\lambda x_3 = 0\\
        \end{cases}
    \]
    and for each permisssible value of \(\lambda\), determine the most general solution.
    \tcblower
    Write this in the form \(Ax=0\) we know 
    \[\begin{bmatrix}
        3 & 1 & -\lambda\\
        4 & -2 & -3\\
        2\lambda & 4 & \lambda
    \end{bmatrix}\]
    and we can use Carmer's rule. To have nontrivial solution, we want \(\det A= 0\) at least, otherwise according to the formula, we'll get \(0\) on the numerator and thus all solution would be trivial. So we want

    \[-6\lambda-6\lambda-16\lambda-4\lambda^2-4\lambda+36 = -4(\lambda^2+8\lambda-9) = -4(\lambda+9)(\lambda-1)=0\]

    this gives \(\lambda= 1, -9\). Nest the general solution. For \(\lambda=1\), after some basic gaussian elimination, we obtain
    \[
        \begin{cases}
            2x_1-x_3 = 0\\
            x_1+x_2 = 0
        \end{cases}
    \]
    So letting \(x_1 = k\), we obtain the solution \(x = [k,-k,2k]\) for \(k\in\bR\).
    
    For \(\lambda=-9\), similarly, we obtain
    \[
        \begin{cases}
            -2x_1-3x_3 = 0\\
            -3x_1+x_2 = 0
        \end{cases}
    \]
    So letting \(x_1 = 3k\), we obtain the solution \(x = [3k,9k,-2k]\) for \(k\in\bR\).

\end{myleftlinebox}

\section*{2}

\begin{myleftlinebox}
    \subsection*{a)} By investigating ranks of relevant matrices, show that the following set of equations possess a one-parameter family of solutions:
    \[
        \begin{cases}
            2x_1 -x_2 -x_3 =2\\
            x_1 +2x_2 +x_3 =2\\
            4x_1 -7x_2 -5x_3 =2
        \end{cases}
    \]
    \subsection*{b)} Determine the general solution
    \tcblower
    \subsection*{a)}
    Write in matrix form \(Ax=b\) where \(A = \begin{bmatrix}
        2 & -1 & -1\\
        1 & 2 & 1\\
        4 & -7 & -5
    \end{bmatrix}\) and the augumented matrix \(A_{aug} = \begin{bmatrix}
        2 & -1 & -1 & 2\\
        1 & 2 & 1 & 2\\
        4 & -7 & -5 & 2
    \end{bmatrix}\).
    Then, to find ranks,

    \begin{equation*}
        rank(A) = rank(\begin{bmatrix}
            0 & -5 & -3 \\
            1 & 2 & 1 \\
            0 & -15 & -9
        \end{bmatrix}) = 2 
    \end{equation*}
    \begin{equation*}
        rank(A_{aug}) = rank(\begin{bmatrix}
            0 & -5 & -3 & -2\\
            1 & 2 & 1 & 2\\
            0 & -15 & -9 & -6
        \end{bmatrix}) = 2
    \end{equation*}
    Thus it's a consistent system and since \(2<3\) thus there should be infinite solutions with the number of independent variables \(3-2=1\), i.e., one-parameter family of solutions.
    \subsection*{b)}
    Start by letting \(x_3 = k\), then from \(-5 x_2-3x_3 = -2\), we have \(x_2 = \dfrac{2-3k}{5}\); next from \(x_1 + 2x_2 +x_3 = 2\), we have \(x_1 = \dfrac{k+6}{5}\) so finally the general solution is \(x = [\frac{k+6}{5},\frac{2-3k}{5},k]\)

\end{myleftlinebox} 

\section*{3}
\begin{myleftlinebox}
    Determine whether the vector \(\{6,1,-6,2\}\) is in the vector space generated by the basis vectors \(\{1,1,-1,1\}\), \(\{-1,0,1,1\}\), \(\{1,-1,-1,0\}\)
    \tcblower
    Write this question in matrix form, we are to solve \(x\) for 
    \[\begin{bmatrix}
        1 & -1 & 1\\
        1 & 0 & -1\\
        -1 & 1 & -1\\
        1 & 1 & 0
    \end{bmatrix} x =  \begin{bmatrix}
        6\\ 1 \\ -6 \\ 2
    \end{bmatrix}\]
    And the solution is \(x=[3,-1,2]\), thus the vector can be represented by the 3 basis vectors.
\end{myleftlinebox}

\section*{4}
\begin{myleftlinebox}
    Find eigen pairs of matrix \(\begin{bmatrix}
        0 & 0 & 2\\
        -1 & 1 & 2\\
        -1 & 0 & 3
    \end{bmatrix}\)
    \tcblower
    First find the characteristic polynomial, denote \(A\) to be the matrix then
    \[\det(\lambda I - A) = \lambda(\lambda-1)(\lambda-3)+2(\lambda-1) = (\lambda-1)^2(\lambda-2)\]
    So the eigenvalues are \(1,2\). For \(\lambda=1\), we solve
    \[(\lambda I - A)x = 0 = \begin{bmatrix}
        1 & 0 & -2 \\
        1 & 0 & -2 \\
        1 & 0 & -2 \\
    \end{bmatrix}x\]
    And this gives the eigenvectors of the form \(x=[2m,k,m]\) for \(m,k \in \bR\). Especially, we have two orthogonal choice, \(x=[2,0,1]\) and \(x = [0,1,0]\).

    For \(\lambda=2\), we solve
    \[(\lambda I - A)x = 0 = \begin{bmatrix}
        2 & 0 & -2\\
        1 & 1 & -2\\
        1 & 0 & =1
    \end{bmatrix}x\]
    And this gives the eigenvectors \(x = [n,n,n]\) for \(n\in \bR\).

\end{myleftlinebox}

\section*{5}
\begin{myleftlinebox}
    Gram-Schmidt orthogonalization for following set of vectors. \([-1,2,0]^\top, [1,1,-1]^\top,[1,-1,1]^\top \).
    \tcblower
    Let the vectors be \(x_i\)s. First we find orthogonal set: \(\hat x_1 = x_1\), \(\hat x_2 = x_2 - (\hat x_1,x_2)x_1/(\hat x_1,\hat x_1) = [1.2,0.6,-1]^\top\), then \(\hat x_3 = x_3 - (\hat x_1,x_3)\hat x_1/(\hat x_1,\hat x_1) - (\hat x_2,x_3)\hat x_2/(\hat x_2,\hat x_2) = [4/7,2/7,6/7]^\top\), then after normalization, we obtain the orthonormal set of vectors: \(u_1 = [-1/\sqrt 5,2/\sqrt 5,0]^\top\), \(u_2 = [6/\sqrt{70},3/\sqrt{70},-5/\sqrt{70}]^\top\), \(u_3 = [4/\sqrt{56},2/\sqrt{56},6/\sqrt{56}]^\top\)
\end{myleftlinebox}

\section*{6}
\begin{myleftlinebox}
    Matrix for the following quadratic form. And determine if it's positive definite, and the shape it makes.
    \[1.5 x_1^2 - x_1x_3 + x_2^2 + 1.5 x_3^2\]
    \tcblower
    As of form \(x^\top A x\), we have \(A = \begin{bmatrix}
        1.5 & 0 & -0.5 \\
        0 & 1 & 0\\
         -0.5 & 0 & 1.5
    \end{bmatrix}\). And its characteristic polynomial is \((\lambda-1)^2(\lambda-2)\) thus all eigenvalues are positive, so positive definite. So it's a Ellipsoid.
\end{myleftlinebox}

\section*{7}
\begin{myleftlinebox}
    Given \(A = \begin{bmatrix}
        3 & -1 & 1\\
        1 & 4 & 0\\
        2 & 1 & -3
    \end{bmatrix}\) and \(B = \begin{bmatrix}
        1 & -3 & 1\\
        2 & 0 & 5\\
        3 & 1 & 2
    \end{bmatrix}\), verify \((AB)^{-1} = B^{-1}A^{-1}\).
    \tcblower
    After some calculations we get, \(A^{-1} = \frac{1}{46}\begin{bmatrix}
        12 & 2 & 4\\
        -3 & 11 & -1\\
        7 & 5 & -13
    \end{bmatrix}\); \(B^{-1} = \frac{1}{36}\begin{bmatrix}
        5 & -7 & 15\\
        -11 & 1 & 3\\
        -2 & 10 & -6
    \end{bmatrix}\); \(AB = \begin{bmatrix}
        4 & -8 & 0\\
        9 & -3 & 21\\
        -5 & -9 & 1
    \end{bmatrix}\) and its inverse \((AB)^{-1} = \frac{1}{828}\begin{bmatrix}
        93 & 4 & -84\\
        -57 & 2 & -42\\
        -48 & 38 & 30
    \end{bmatrix}\) and \( B^{-1}A^{-1} = \frac{1}{1656}\begin{bmatrix}
        186 & 8 & -168\\
        -114 & 4 & -84\\
        -96 & 76& 60
    \end{bmatrix} = (AB)^{-1} \)
\end{myleftlinebox}



%\newpage

%\bibliographystyle{acm}
%\bibliography{D:/Papers.bib}

%\newpage
%\appendix
%\section{Appendix}

\end{document}